---
title: "Loop_second_model"
author: "Gemma Izquierdo"
date: "2026-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This .Rmd contains the code to run the Kalman filter across the 6 trials and 5 channels for each combination of p and fs. The parameters are stored in lists, which will be used in the analysis.

```{r}
library(signal)
library(dlm)
library(readxl)
ordenSensores <- read_excel("ordenSensores.xls", col_names = FALSE) # read .xls with the names of the electrodes
names(ordenSensores)<-c("electrode")

EEGtrial <- readRDS("matrices.rds") # read EEG data
trials <- c(1,2,4,5,21,223) # vector with trials to consider
channels_index <- c(24, 27,  8, 57, 39)
source("Functions.R")
```

This .rmd contains the loops to store the parameters estimated for each set of signals in function of the sampling frequency (fs) and number of neighbors (p-1).

# AR(1)
# fs=250
## p=1

```{r}
p <- 1 
parameters_trials_notfull_p1_phi_250 <- list() # list to store the parameters estimated with p = 1 and sampling frequency = 250 Hz
chan_names<-ordenSensores$electrode 
npar<-(p^2+p+2)/2+1 # number of parameters to estimate

for (trial in trials){

  parameters_trials_notfull_p1_phi_250[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]]) # EEG data from trial selected

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE) # computation of correlation matrix for channel ch and ordering in function of the correlation value (reference channel ch is the first one)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }

    ds <- 4 # factor by which the signal is downsampled
    y <- eeg_data[, idx[1:p]] # reference signal + (p-1) most correlated signals
    y <- downsample(y, ds)

    # Maximum likelihood estimation
    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar), # initialization of parameters
      build = buildSignalnotfullphineq0_p1,  # function to build the filter
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p1_phi_250[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=250
## p=3

```{r}
trials <- c(1,2,4,5,21,223)
p <- 3
parameters_trials_notfull_p3_phi_250 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p3_phi_250[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 4
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p3_phi_250[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=250
## p=5

```{r}
trials <- c(1,2,4,5,21,223)
p <- 5
parameters_trials_notfull_p5_phi_250 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p5_phi_250[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 4
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p5_phi_250[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=250
## p=5

```{r}
trials <- c(1,2,4,5,21,223)
p <- 7
parameters_trials_notfull_p7_phi_250 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p7_phi_250[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 4
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p7_phi_250[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=500
## p=1

```{r}
trials <- c(1,2,4,5,21,223)
p <- 1
parameters_trials_notfull_p1_phi_500 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p1_phi_500[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }

    ds <- 2
    y <- eeg_data[, idx[1:p]]
    y <- downsample(y, ds)
    #colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0_p1,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p1_phi_500[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=500
## p=3

```{r}
library(signal)
library(dlm)

trials <- c(1,2,4,5,21,223)
p <- 3
parameters_trials_notfull_p3_phi_500 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p3_phi_500[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 2
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p3_phi_500[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=500
## p=5

```{r}
trials <- c(1,2,4,5,21,223)
p <- 5
parameters_trials_notfull_p5_phi_500 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p5_phi_500[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 2
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p5_phi_500[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=500
## p=7

```{r}
trials <- c(1,2,4,5,21,223)
p <- 7
parameters_trials_notfull_p7_phi_500 <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p7_phi_500[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    ds <- 2
    y <- eeg_data[, idx[1:p]]
    y <- apply(y, 2, function(x) downsample(x, ds))
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p7_phi_500[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=1000
## p=1

```{r}
trials <- c(1,2,4,5,21,223)
p <- 1
parameters_trials_notfull_p1_phi <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p1_phi[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }

    y <- eeg_data[, idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0_p1,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p1_phi[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=1000
## p=3

```{r}
library(signal)
library(dlm)

trials <- c(1,2,4,5,21,223)
p <- 3
parameters_trials_notfull_p3_phi <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p3_phi[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    y <- eeg_data[, idx[1:p]]
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p3_phi[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=1000
## p=5

```{r}
trials <- c(1,2,4,5,21,223)
p <- 5
parameters_trials_notfull_p5_phi <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p5_phi[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    y <- eeg_data[, idx[1:p]]
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p5_phi[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```

# AR(1)
# fs=1000
## p=7

```{r}
trials <- c(1,2,4,5,21,223)
p <- 7
parameters_trials_notfull_p7_phi <- list()
chan_names<-ordenSensores$electrode
npar<-(p^2+p+2)/2+1

for (trial in trials){

  parameters_trials_notfull_p7_phi[[as.character(trial)]] <- list()

  eeg_data <- t(EEGtrial[[trial]])

  for (ch in channels_index){

    cat("\nTrial:", trial, "Channel:", ch, "-", chan_names[ch])

    idx <- order(cor(eeg_data)[,ch], decreasing = TRUE)

    for (i in 1:p){
      cat("\n    Channel:", idx[i])
    }
    
    y <- eeg_data[, idx[1:p]]
    colnames(y) <- chan_names[idx[1:p]]

    fitSignal <- dlmMLE(
      y,
      parm = rep(0, npar),
      build = buildSignalnotfullphineq0,
      hessian = TRUE,
      control = list(maxit = 500)
    )

    parameters_trials_notfull_p7_phi[[as.character(trial)]][[ch]] <- fitSignal$par
  }
}
```


