---
title: "Q_Q_plots"
author: "Gemma Izquierdo"
date: "2026-09-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This .Rmd shows the code to compute the Q-Q plots once the parameters are estimated, as well as the Shapiro-Wilk test.

Loading of parameters estimations.

```{r}
library(ggplot2)
library(dplyr)

source("Functions.R")

EEGtrial <- readRDS("matrices.rds")

parameters_trials_notfull_p3_phi<-readRDS("parameters_trials_notfull_p3_phi.rds")
parameters_trials_notfull_p5_phi<-readRDS("parameters_trials_notfull_p5_phi.rds")
parameters_trials_notfull_p7_phi<-readRDS("parameters_trials_notfull_p7_phi.rds")
parameters_trials_notfull_p1_phi<-readRDS("parameters_trials_notfull_p1_phi.rds")

parameters_trials_notfull_p3_phi_500<-readRDS("parameters_trials_notfull_p3_phi_500.rds")
parameters_trials_notfull_p5_phi_500<-readRDS("parameters_trials_notfull_p5_phi_500.rds")
parameters_trials_notfull_p7_phi_500<-readRDS("parameters_trials_notfull_p7_phi_500.rds")
parameters_trials_notfull_p1_phi_500<-readRDS("parameters_trials_notfull_p1_phi_500.rds")

parameters_trials_notfull_p3_phi_250<-readRDS("parameters_trials_notfull_p3_phi_250.rds")
parameters_trials_notfull_p5_phi_250<-readRDS("parameters_trials_notfull_p5_phi_250.rds")
parameters_trials_notfull_p7_phi_250<-readRDS("parameters_trials_notfull_p7_phi_250.rds")
parameters_trials_notfull_p1_phi_250<-readRDS("parameters_trials_notfull_p1_phi_250.rds")

channels_index <- c(24, 27,  8, 57, 39)
trials <-  c(1, 2, 4, 5, 21, 223)
```

Computation of the residuals and Shapiro-Wilk test for p = 1.

```{r}
library(dplyr)
library(ggplot2)

p <- 1

# List of models with different sampling frequency (fs) and p=1
models <- list(
  list(
    ds = 1,
    pars = parameters_trials_notfull_p1_phi,
    title = "1000 Hz"
  ),
  list(
    ds = 2,
    pars = parameters_trials_notfull_p1_phi_500,
    title = "500 Hz"
  ),
  list(
    ds = 4,
    pars = parameters_trials_notfull_p1_phi_250,
    title = "250 Hz"
  )
)

residuals_results <- list()

# Loop across 6 trials
for (trial_n in seq_along(trials)) {
  
  eeg_data <- t(EEGtrial[[trials[trial_n]]])
  cor_mat <- cor(eeg_data) # correlation matrix
  
  # Loop across 5 channels
  for (ch in channels_index) {
    
    idx <- order(cor_mat[, ch], decreasing = TRUE) # Ordering in decreasing order (first one is the reference channel)
    
    for (m in models) {
      
      y <- downsample(eeg_data[, idx[1:p]], m$ds) # Downsample
      
      par <- m$pars[[trial_n]][[ch]]
      
      # Build filtered signal
      dlmM1 <- buildSignalnotfullphineq0_p1(par)
      # Smooth signal
      eegSmo <- dlmSmooth(y, dlmM1)
      
      # Compute residuals
      signal <- dropFirst(eegSmo$s)
      residuals <- y - signal
      
      # Store residuals and Shapiro-Wilk test results
      residuals_results[[length(residuals_results) + 1]] <- data.frame(
        trial = trials[trial_n],
        channel = ordenSensores$electrode[ch],
        fs = m$title,
        residual = residuals,
        shapirotest = shapiro.test(residuals)$p.value
      )
    }
  }
}

residuals_df <- bind_rows(residuals_results)
```

Q-Q plot for p = 1 in function of sampling frequency (fs). Inside the code, you must choose the fs that you want to plot. The example is given for fs = 250 Hz.

```{r}
# Select the fs to be plotted (here 250 Hz)
residuals_plot <- residuals_df %>%
  dplyr::filter(fs == "250 Hz")

# Highlight channels that fail the Shapiro-Wilk test
highlight <- residuals_plot %>%
  dplyr::filter(shapirotest < 0.05) %>%
  mutate(id = paste(trial, channel)) %>%
  pull(id)

residuals_plot$highlight <- paste(
  residuals_plot$trial,
  residuals_plot$channel
) %in% highlight

p1_250 <- ggplot(residuals_plot, aes(sample = residual)) +
  stat_qq(shape = 1) +
  geom_qq(
    data = residuals_plot[residuals_plot$highlight, ],
    shape = 21,
    colour = "blue",
    size = 2
  ) +
  stat_qq_line(colour = "red") +
  facet_grid(
    channel ~ fs + trial,
    scales = "free"
  ) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Q-Q plot p=1"
  ) +
  theme_bw()

p1_250
```

Computation of the residuals and Shapiro-Wilk test for p > 1.


```{r}
library(dplyr)
library(ggplot2)

p <- 3

models <- list(
  list(
    ds = 1,
    pars = parameters_trials_notfull_p3_phi,
    title = "1000 Hz"
  ),
  list(
    ds = 2,
    pars = parameters_trials_notfull_p3_phi_500,
    title = "500 Hz"
  ),
  list(
    ds = 4,
    pars = parameters_trials_notfull_p3_phi_250,
    title = "250 Hz"
  )
)

residuals_results <- list()

for (trial_n in seq_along(trials)) {
  
  eeg_data <- t(EEGtrial[[trials[trial_n]]])
  cor_mat <- cor(eeg_data)
  
  for (ch in channels_index) {
    
    idx <- order(cor_mat[, ch], decreasing = TRUE)
    
    for (m in models) {
      
              
      y <- eeg_data[, idx[1:p]]

      y <- apply(
        y,
        2,
        downsample,
        m$ds
      )
      
      par <- m$pars[[trial_n]][[ch]]
      
      # Build filtered signal
      dlmM1 <- buildSignalnotfullphineq0(par)
      # Smooth signal
      eegSmo <- dlmSmooth(y, dlmM1)
      
      # Compute residuals
      signal <- dropFirst(eegSmo$s)
      residuals <- y[,1] - signal
      
      residuals_results[[length(residuals_results) + 1]] <- data.frame(
        trial = trials[trial_n],
        channel = ordenSensores$electrode[ch],
        fs = m$title,
        residual = residuals,
        shapirotest = shapiro.test(residuals)$p.value
      )
    }
  }
}

residuals_df <- bind_rows(residuals_results)
```

Q-Q plot for p = 1 in function of sampling frequency (fs). Inside the code, you must choose the fs that you want to plot. The example is given for fs = 250 Hz.

```{r}
# Select the fs to be plotted (here 500 Hz)
residuals_plot <- residuals_df %>%
  dplyr::filter(fs == "250 Hz")

# Highlight channels that fail the Shapiro-Wilk test
highlight <- residuals_plot %>%
  dplyr::filter(shapirotest < 0.05) %>%
  mutate(id = paste(trial, channel)) %>%
  pull(id)

residuals_plot$highlight <- paste(
  residuals_plot$trial,
  residuals_plot$channel
) %in% highlight

p3_250 <- ggplot(residuals_plot, aes(sample = residual)) +
  stat_qq(shape = 1) +
  geom_qq(
    data = residuals_plot[residuals_plot$highlight, ],
    shape = 21,
    colour = "blue",
    size = 2
  ) +
  stat_qq_line(colour = "red") +
  facet_grid(
    trial ~ fs + channel,
    scales = "free"
  ) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Q-Q plot p=3"
  ) +
  theme_bw()

p3_250
```
